---
title: "R for Biologists"
author: "Saville"
format: html
editor: visual
---

# Intro to R for Biologists

This is a basic intro to R and the Tidyverse for biologists. The link for this brief course is here: [Intro R for Biologists](https://melbournebioinformatics.github.io/r-intro-biologists/intro_r_biologists.html)

It uses the tidyverse, so we need to install the tidyverse package

```{r}
install.packages("tidyverse")
```

As well as load the tidyverse library

```{r}
library(tidyverse)
```

It uses the following RNA seq data

### RNA seq dataset

The data used are from the paper by [Fu et al. 2015](https://www.ncbi.nlm.nih.gov/pubmed/25730472), GEO code GSE60450. This study examined expression in basal and luminal cells from mice at different stages (virgin, pregnant and lactating). There are 2 samples per group and 6 groups, 12 samples in total.

This is the same data used in the Galaxy RNA seq tutorial (also written by Maria Doyle - and others)

[galaxy RNA seq tutorial](https://training.galaxyproject.org/training-material/topics/transcriptomics/tutorials/rna-seq-reads-to-counts/tutorial.html)

let's see if we can read the data from the github page

```{r}
library(readr)
```

```{r}
counts <- read_csv(url("https://raw.githubusercontent.com/melbournebioinformatics/r-intro-biologists/master/data/GSE60450_GeneLevel_Normalized(CPM.and.TMM)_data.csv"))
```

We also need to read in the meta data that describes the samples.

```{r}
sample_info <- read_csv(url("https://raw.githubusercontent.com/melbournebioinformatics/r-intro-biologists/master/data/GSE60450_filtered_metadata.csv"))
```

We can also read these files from the data folder. I manually downloded these from the github page and uploaded them to this project.

Uncomment and run the commands below if the above doesn't work.

```{r}
### can also read from data contained in the dat folder of this project.  Uncomment the two lines below to import the data locally

#counts <- read_csv("data/GSE60450_GeneLevel_Normalized(CPM.and.TMM)_data.csv")

#sample_info <- read_csv("data/GSE60450_filtered_metadata.csv")

```
